alternativeRSA

### 2016-10-02 Paper/Paper.tex
sanitize

\begin{abstract}
Successful communication regularly requires listeners to make pragmatic
inferences --- enrichments beyond the literal meaning of a speaker's
utterance. For example, when interpreting a sentence such as ``Alice ate
some of the cookies,'' listeners routinely infer that Alice did not eat
all of them. A Gricean account of this phenomena assumes the presence of
alternatives (like ``all of the cookies'') with varying degrees of
informativity, but it remains an open question precisely what these
alternatives are. To address this question, we collect empirical
measurements of speaker and listener judgments about varying sets of
alternatives across a range of scales and use these as inputs to a
computational model of pragmatic inference. This approach allows us to
test hypotheses about how well different sets of alternatives predict
pragmatic judgments by people. Our findings suggest that comprehenders
likely consider a broader set of alternatives beyond those logically
entailed by the initial message.

@@

How much of what we mean comes from the words that go unsaid? As
listeners, our ability to make precise inferences about a speaker's
intended meaning in context is indispensable to successful
communication. For example, listeners commonly enrich the meaning of the
scalar item \emph{some} to \emph{some but not all} in sentences like
``Alice ate some of the cookies'' (Grice, 1975; Levinson, 2000). These
inferences, called \emph{scalar implicatures}, have been an important
test case for understanding pragmatic inferences more generally. A
Gricean account of this phenomenon assumes listeners reason about a
speaker's intended meaning by incorporating knowledge about A)
alternative scalar items a speaker could have used (such as \emph{all})
and B) the relative informativity of using such alternatives (Grice,
1975). According to this account, a listener will infer that the speaker
must have intended that Alice did not eat all the cookies because it
would have been underinformative to use the descriptor \emph{some} when
the alternative \emph{all} could have been used just as easily.

But what are the alternatives that should be considered in the
implicature computation more generally? Under classic accounts,
listeners consider only those words whose meanings entail the actual
message (Horn, 1972), and these alternatives enter into conventionalized
or semi-conventionalized scales (Levinson, 2000). For example, because
\emph{all} entails \emph{some}, and hence is a ``stronger'' meaning,
\emph{all} should be considered as an alternative to \emph{some} in
implicatures. Similar scales exist for non-quantifier scales, e.g.
\emph{loved} entails \emph{liked} (and hence ``I liked the movie''
implicates that I didn't love it).

Recent empirical evidence has called into question whether entailment
scales are all that is necessary for understanding scalar implicature.
For example, Degen \& Tanenhaus (2015) demonstrated that the scalar item
\emph{some} was judged less appropriate when exact numbers were seen as
viable alternatives. And in a different paradigm, {van Tiel} (2014)
found converging evidence that \emph{some} was judged to be atypical for
small quantities. These data provide indirect evidence that people may
actually consider a broader set of alternatives when computing scalar
implicatures. Since \emph{some} is logically true of sets with one or
two members, these authors argued that the presence of salient
alternatives (the words \emph{one} and \emph{two}, for example) reduced
the felicity of \emph{some} via a pragmatic inference.

@@

Franke (2014) used an RSA-style model to assess what alternatives a
speaker would need to consider in order to produce the
typicality/felicity ratings reported by Degen \& Tanenhaus (2015) and
{van Tiel} (2014) for the scale \emph{some}/\emph{all}. In order to do
this, Franke (2014)'s model assigned weights to a set of alternative
numerical expressions. Surprisingly, along with weighting \emph{one}
highly (a conclusion that was supported by the empirical work), the
best-fitting model assigned substantial weight to \emph{none} as an
alternative to \emph{some}. This finding was especially surprising
considering the emphasis of standard theories on scalar items that stand
in entailment relationships with one another (e.g. \emph{one} entails
\emph{some} even if it is not classically considered to be part of the
scale).

We pick up where these previous studies left off, considering the set of
alternatives for implicature using the RSA model. To gain empirical
traction on this issue, we broaden the set of scales we consider. Our
inspiration for this move comes from work by {van Tiel}, Van Miltenburg,
Zevakhina, \& Geurts (2014), who examined a phenomenon that they dubbed
``scalar diversity,'' namely the substantial difference in the strength
of scalar implicature across a variety of scalar pairs (e.g.,
\emph{liked} / \emph{loved}, \emph{palatable} / \emph{delicious}).
Making use of this diversity allows us to investigate the ways that
different alternative sets give rise to implicatures of different
strengths across scales.

@@

\caption[Combined counts for participant-generated alternatives for liked vs]{Combined counts for participant-generated alternatives for liked vs. loved in Experiment 1.}\label{fig:exp2_altsPlot_likedLoved}
\end{figure}
\end{CodeChunk}

@@

To examine the effect different alternative sets have on implicature, we
needed a way of expanding alternative sets beyond entailment pairs. We
addressed this issue by adopting a modified cloze task to elicit
alternatives empirically. This design was inspired by Experiment 3 of
{van Tiel} et al. (2014).

@@

On each trial participants were presented with a target scalar item from
five scales, selected from the entailment items used in {van Tiel} et
al. (2014). These were embedded in a sentence such as, ``In a recent
restaurant review someone said they thought the food was \_\_\_\_," with
a target scalar presented in the blank. Participants were then asked to
generate plausible alternatives by responding to the question, ``If
they'd felt differently about the food, what other words could they have
used instead of \_\_\_\_?" They were prompted to generate three unique
alternatives.

@@

@@

\caption[(Left) A trial from Experiment 2 (literal listener) with the target scalar ``liked]{(Left) A trial from Experiment 2 (literal listener) with the target scalar ``liked." (Right) A trial from Experiment 3 (pragmatic listener) with the target scalar ``liked."}\label{fig:stimuli_exp1}
\end{figure*}
\end{CodeChunk}

\section{Experiment 2: Literal listener
task}\label{experiment-2-literal-listener-task}

Experiment 2 was conducted to approximate literal listener
semantics---\(p_{L_0}(m \mid u)\) in Equation (3)---for the same five
pairs of scalar items used in Experiment 1. We included three
conditions: two alternatives (``entailment''), four alternatives, and
five alternatives. The two alternatives condition makes a test of the
hypothesis that the two members of the classic Horn (entailment) scale
(Horn, 1972) are the only alternatives necessary to predict the strength
of listeners' pragmatic inference. The four and five alternative
conditions then add successively more alternatives to test whether
including a larger number of alternatives will increase model
fit.\footnote{Note that alternatives in the four and five alternative conditions were chosen on the basis of Experiment 1, which was run chronologically after the two alternative condition; all literal listener experiments are grouped together for simplicity in reporting.}
A secondary goal of Experiment 2 is to test whether the set of
alternatives queried during literal semantic elicitation impacts literal
semantic judgments. If it does we should see differences in these
judgments for shared items between conditions

@@

Conditions were run sequentially. In each condition we recruited 30
participants from AMT. In the two alternatives condition, 16
participants were excluded for either failing to pass two training
trials or because they were not native English speakers, leaving a total
sample of 14
participants.\footnote{The majority of respondent data excluded from the two alternatives condition was caused by failure to pass training trials. We believe the task may have been too difficult for most respondents and made adjustments to the training trials in later conditions.}
In the four alternatives condition, 7 participants were excluded for
either failing to pass two training trials or not being native English
speakers, leaving a total sample of 23 participants. In the five
alternatives condition, 3 participants were excluded for either failing
to pass two training trials or not being native English speakers,
leaving a total sample of 27 participants.

@@

The two alternatives condition only included entailment pairs taken from
{van Tiel} et al. (2014). The four alternatives condition included
entailment pairs plus the top two alternatives generated for each scalar
family by participants in Experiment 1. The five alternatives condition
included the four previous items plus one more neutral item chosen from
alternatives generated in Experiment 1.

@@

\section{Experiment 3: Pragmatic
Listener}\label{experiment-3-pragmatic-listener}

@@

\subsection{Participants}\label{participants-2}

@@

\subsection{Procedure}\label{procedure}

Participants were presented with a one-sentence prompt such as ``Someone
said they thought the food was \_\_\_\_\_.'' with a target scalar item
in the blank. Participants were then asked to generate a star rating
representing the rating they thought the reviewer likely gave. Each
participant was presented with all scalar items in a random order. The
experimental setup is shown in Figure \ref{fig:stimuli_exp1}, right.

@@

\caption[Pragmatic listener judgments from Experiment 3]{Pragmatic listener judgments from Experiment 3. Proportion of participants generating a star rating is shown on the vertical axis, with the horizontal axis showing number of stars on which the utterance was judged. Line type denotes condition, and colors indicate the particular scalar items. Each panel shows one scalar pair, with only standard entailment pairs shown here for simplicity in reporting.}\label{fig:exp2Plots}
\end{figure*}
\end{CodeChunk}

Figure \ref{fig:exp2Plots} plots pragmatic listener judgment
distributions for ``weak'' / ``strong'' scalar pairs (e.g.
\emph{good}/\emph{excellent}). Several trends are visible. First, in
each scale participants generated implicatures. They were substantially
less likely to assign high star ratings to weaker scalar terms, despite
the literal semantic compatibility of those terms with those states
shown in Experiment 2. Second, the difference between strong and weak
scalar items varied considerably across scales, consistent with previous
work ({van Tiel} et al., 2014).

@@

\section{Model}\label{model}

Using literal listener data from Experiment 2, we conducted a set of
simulations with the RSA model. Each simulation kept the model constant,
fitting the choice parameter \(\alpha\) as a free parameter, but used a
set of alternatives to specify the scale over which predictions were
computed. We considered four different alternative sets, with empirical
measurements corresponding to those shown in Figure
\ref{fig:allScalesTable}: 1) the two alternatives in the classic
entailment scales, 2) those two alternatives with the addition of a
generic negative alternative, 3) the expanded set of four alternatives,
and 4) the expanded set of five alternatives. Literal semantics for the
generic negative alternative served as a baseline ``none``-style
semantics in which the scalar item was only compatible with 1 star.

Model fit with human judgments was significantly improved by the
inclusion of alternatives beyond the entailment items (Table 1). The two
alternatives model contained only entailment items, which, under classic
accounts, should be sufficient to generate implicature, but fit to data
was quite poor with these items. The addition of a generic negative
element produced some gains, but much higher performance was found when
we included four and five alternatives, with the alternatives derived
empirically for the specific scale we used. An example fit for the five
alternatives model is shown in Figure \ref{fig:fiveAltsScatter}.

\begin{table}[ht]
\centering
\begin{tabular}{lrrr}
  \hline
Model & $\alpha$ & Two alts & Four alts \\ 
  \hline
Two alts &   9 & 0.54 & 0.57 \\ 
  Two alts + generic negative  &   6 & 0.62 & 0.66 \\ 
  Four alts &   4 & 0.84 & 0.90 \\ 
  Five alts &   4 & 0.86 & 0.90 \\ 
   \hline
\end{tabular}
\caption{Model performance with fitted alpha levels. Model fit assessed through correlation with human judgments in the two conditions of Experiment 3.} 
\end{table}

@@

In our current work, we used the rational speech act framework to
investigate the set of alternatives that best allowed the model to
predict pragmatic judgments across a series of different scales. We
found that the best predictions came when a range of scale-dependent
negative and neutral alternatives were added to the implicature
computation, suggesting the importance of considering non-entailment
alternatives. This work builds on previous investigations, reinforcing
the claim that negative alternatives are critical for understanding
implicature (Franke, 2014), and replicates and extends findings that
different lexical scales produce strikingly different patterns of
inference ({van Tiel} et al., 2014).

@@

Degen, J., \& Tanenhaus, M. K. (2015). Processing scalar implicature: A
constraint-based approach. \emph{Cognitive Science}, \emph{39}(4),
667--710.

Frank, M., \& Goodman, N. D. (2012). Predicting pragmatic reasoning in
language games. \emph{Science}, \emph{336}(6084), 998.

Franke, M. (2014). Typical use of quantifiers: A probabilistic speaker
model. In \emph{Proceedings of the 36th annual conference of the
cognitive science society} (pp. 487--492).

Goodman, N. D., \& Stuhlmuller, A. (2013). Knowledge and implicature:
Modeling language understanding as social cognition. \emph{Topics in
Cognitive Science}, \emph{5}(1), 173--184.

Grice, H. P. (1975). Logic and conversation. In P. Cole \& J. Morgan
(Eds.), \emph{Syntax and semantics} (Vol. 3). New York: Academic Press.

Horn, L. R. (1972). \emph{On the semantic properties of logical
operators.} (PhD thesis). University of California, Los Angeles.

Levinson, S. C. (2000). \emph{Presumptive meanings: The theory of
generalized conversational implicature}. MIT Press.

Lewis, D. (1969). \emph{Convention: A philosophical study}. John Wiley
\& Sons.

Sperber, D., \& Wilson, D. (1995). \emph{Relevance: Communication and
cognition} (2nd ed.). Oxford, UK: Blackwell.

{van Tiel}, B. J. M. (2014). Quantity matters: Implicatures, typicality,
and truth.

{van Tiel}, B. J. M., Van Miltenburg, E., Zevakhina, N., \& Geurts, B.
(2014). Scalar diversity. \emph{Journal of Semantics}, ffu017.

>>>>

\begin{abstract}
Successful communication regularly requires listeners to make pragmatic
inferences --- enrichments beyond the literal meaning of a speaker's
utterance. For example, when interpreting a sentence such as ``Alice ate
some of the cookies,'' listeners routinely infer that Alice did not eat
all of them. A Gricean account of this phenomenon assumes the presence
of alternatives (like ``all of the cookies'') with varying degrees of
informativity, but it remains an open question precisely what these
alternatives are. To address this question, we collect empirical
measurements of speaker and listener judgments about varying sets of
alternatives across a range of scales and use these as inputs to a
computational model of pragmatic inference. This approach allows us to
test hypotheses about how well different sets of alternatives predict
pragmatic judgments by people. Our findings suggest that comprehenders
likely consider a broader set of alternatives beyond those logically
entailed by the initial message.

@@

How much of what we mean comes from the words that go unsaid? As
listeners, our ability to make precise inferences about a speaker's
intended meaning in context is indispensable to successful
communication. For example, listeners commonly enrich the meaning of the
word \emph{some} to \emph{some but not all} in sentences like ``Alice
ate some of the cookies'' (Grice, 1975; Levinson, 2000). These
inferences, called \emph{scalar implicatures}, have been an important
test case for understanding pragmatic inferences more generally. A
Gricean account of this phenomenon assumes listeners reason about a
speaker's intended meaning by incorporating knowledge about A)
alternative scalar items a speaker could have used (such as \emph{all})
and B) the relative informativity of using such alternatives (Grice,
1975). According to this account, a listener will infer that the speaker
intended to convey that Alice did not eat all the cookies because it
would have been underinformative to use the descriptor \emph{some} when
the alternative \emph{all} could have been used just as easily.

But what are the alternatives that should be considered in the
implicature computation more generally? Under classic accounts,
listeners consider only those words whose meanings entail the actual
message (Horn, 1972), and these alternatives enter into conventionalized
or semi-conventionalized scales (Levinson, 2000). For example, because
\emph{all} entails \emph{some}, and hence is a ``stronger'' meaning,
\emph{all} should be considered as an alternative to \emph{some} in
implicatures. Similar scales exist for non-quantifier scales, e.g.
\emph{loved} entails \emph{liked} (hence ``I liked the movie''
implicates that I didn't love it).

Recent empirical evidence has called into question whether entailment
scales are all that is necessary for understanding scalar implicature.
For example, Degen \& Tanenhaus (2015) demonstrated that the scalar item
\emph{some} was judged less appropriate when exact numbers were seen as
viable alternatives. And in a different paradigm, van Tiel (2014) found
converging evidence that \emph{some} was judged to be atypical for small
quantities. These data provide indirect evidence that people may
actually consider a broader set of alternatives when computing scalar
implicatures. Since \emph{some} is logically true of sets with one or
two members, these authors argued that the presence of salient
alternatives (the words \emph{one} and \emph{two}, for example) reduced
the felicity of \emph{some} via a pragmatic inference.

@@

Franke (2014) used an RSA-style model to assess what alternatives a
speaker would need to consider in order to produce the
typicality/felicity ratings reported by Degen \& Tanenhaus (2015) and
van Tiel (2014) for the scale \emph{some}/\emph{all}. In order to do
this, Franke (2014)'s model assigned weights to a set of alternative
numerical expressions. Surprisingly, along with weighting \emph{one}
highly (a conclusion that was supported by the empirical work), the
best-fitting model assigned substantial weight to \emph{none} as an
alternative to \emph{some}. This finding was especially surprising
considering the emphasis of standard theories on scalar items that stand
in entailment relationships with one another (e.g. \emph{one} entails
\emph{some} even if it is not classically considered to be part of the
scale).

We pick up where these previous studies left off, considering the set of
alternatives for implicature using the RSA model. To gain empirical
traction on this issue, we broaden the set of scales we consider. Our
inspiration for this move comes from work by van Tiel, Van Miltenburg,
Zevakhina, \& Geurts (2014), who examined a phenomenon that they dubbed
``scalar diversity,'' namely the substantial difference in the strength
of scalar implicature across a variety of scalar pairs (e.g.,
\emph{liked} / \emph{loved}, \emph{palatable} / \emph{delicious}).
Making use of this diversity allows us to investigate the ways that
different alternative sets give rise to implicatures of different
strengths across scales.

@@

\caption[Combined counts for participant-generated alternatives for liked and loved in Experiment 1]{Combined counts for participant-generated alternatives for liked and loved in Experiment 1.}\label{fig:exp2_altsPlot_likedLoved}
\end{figure}
\end{CodeChunk}

@@

To examine the effect different alternative sets have on implicature, we
needed a way of expanding alternative sets beyond entailment pairs. We
addressed this issue by adopting a modified cloze task to elicit
alternatives empirically. This design was inspired by Experiment 3 of
van Tiel et al. (2014).

@@

On each trial participants were presented with a target scalar item from
five scales, selected from the entailment items used in van Tiel et al.
(2014). These were embedded in a sentence such as, ``In a recent
restaurant review someone said they thought the food was \_\_\_\_," with
a target scalar presented in the blank. Participants were then asked to
generate three plausible alternatives by responding to the question,
``If they'd felt differently about the food, what other words could they
have used instead of \_\_\_\_?" Participants saw each scalar item in
random order. In total, each participant saw ten target scalar items
(each entailment pair from the five scales selected from van Tiel et
al., 2014).

@@

\section{Experiment 2: Literal Listener
Task}\label{experiment-2-literal-listener-task}

Experiment 2 was conducted to approximate literal listener
semantics---\(p_{L_0}(m \mid u)\) in Equation (3)---for the same five
pairs of scalar items used in Experiment 1. We included three
conditions: two alternatives (``entailment''), four alternatives, and
five alternatives. The two alternatives condition makes a test of the
hypothesis that the two members of the classic Horn (entailment) scale
(Horn, 1972) are the only alternatives necessary to predict the strength
of listeners' pragmatic inference. The four and five alternative
conditions then add successively more alternatives to test whether
including a larger number of alternatives will increase model
fit.\footnote{Note that alternatives in the four and five alternative conditions were chosen on the basis of Experiment 1, which was run chronologically after the two alternative condition; all literal listener experiments are grouped together for simplicity in reporting.}
A secondary goal of Experiment 2 is to test whether the set of
alternatives queried during literal semantic elicitation impacts literal
semantic judgments. If it does we should see differences in these
judgments for shared items between conditions.

@@

\caption[(Left) A trial from Experiment 2 (literal listener) with the target scalar 'liked']{(Left) A trial from Experiment 2 (literal listener) with the target scalar 'liked'. (Right) A trial from Experiment 3 (pragmatic listener) with the target scalar 'liked'.}\label{fig:stimuli_exp1}
\end{figure*}
\end{CodeChunk}

@@

Conditions were run sequentially. In each condition we recruited 30
participants from AMT. In the two alternatives condition, 16
participants were excluded for either failing to pass two training
trials or because they were not native English speakers, leaving a total
sample of 14
participants.\footnote{The majority of respondent data excluded from the two alternatives condition was caused by failure to pass training trials. We believe the task may have been too difficult for most respondents who may have been confused by the question wording. Training trials in later conditions included clearer language.}
In the four alternatives condition, 7 participants were excluded for
either failing to pass two training trials or not being native English
speakers, leaving a total sample of 23 participants. In the five
alternatives condition, 3 participants were excluded for either failing
to pass two training trials or not being native English speakers,
leaving a total sample of 27 participants.

@@

The two alternatives condition only included entailment pairs taken from
van Tiel et al. (2014) for a total of 50 trials for each participant.
The four alternatives condition included entailment pairs plus the top
two alternatives generated for each scalar family by participants in
Experiment 1 (100 trials per participant). The five alternatives
condition included the four previous items plus one more neutral item
chosen from alternatives generated in Experiment 1 (125 trials per
participant).

@@

\section{Experiment 3: Pragmatic Listener
Task}\label{experiment-3-pragmatic-listener-task}

@@

\subsection{Methods}\label{methods-2}

\subsubsection{Participants}\label{participants-2}

@@

\subsubsection{Design and Procedure}\label{design-and-procedure-2}

Participants were presented with a one-sentence prompt such as ``Someone
said they thought the food was \_\_\_\_\_.'' with a target scalar item
in the blank. Participants were then asked to generate a star rating
representing the rating they thought the reviewer likely gave. Each
participant was presented with all scalar items in a random order.
Participants in the two alternatives condition gave a total of 10
pragmatic judgments. Participants in the four alternatives condition
gave a total of 20 pragmatic judgments. The experimental setup is shown
in Figure \ref{fig:stimuli_exp1}, right.

@@

\caption[Pragmatic listener judgments from Experiment 3]{Pragmatic listener judgments from Experiment 3. Vertical axis shows proportion of participants generating a star rating. Horizontal axis shows number of stars on which the utterance was judged. Line type denotes condition and colors indicate the particular scalar items. Each panel shows one entailment scalar pair.}\label{fig:exp2Plots}
\end{figure*}
\end{CodeChunk}

Figure \ref{fig:exp2Plots} plots pragmatic listener judgment
distributions for ``weak'' / ``strong'' scalar pairs (e.g. \emph{good} /
\emph{excellent}). Several trends are visible. First, in each scale
participants generated implicatures. They were substantially less likely
to assign high star ratings to weaker scalar terms, despite the literal
semantic compatibility of those terms with those states shown in
Experiment 2. Second, the difference between strong and weak scalar
items varied considerably across scales, consistent with previous work
(van Tiel et al., 2014).

@@

\section{Model Simulations}\label{model-simulations}

Using literal listener data from Experiment 2, we conducted a set of
simulations with the RSA model. Each simulation kept the model constant,
fitting the choice parameter \(\alpha\) as a free parameter, but used a
set of alternatives to specify the scale over which predictions were
computed. We considered four different alternative sets, with empirical
measurements corresponding to those shown in Figure
\ref{fig:allScalesTable}: 1) the two alternatives in the classic
entailment scales, 2) those two alternatives with the addition of a
generic negative alternative, 3) the expanded set of four alternatives,
and 4) the expanded set of five alternatives. Literal semantics for the
generic negative alternative served as a baseline ``none" semantics in
which the scalar item was only compatible with 1 star.

Model fit with human judgments was significantly improved by the
inclusion of alternatives beyond the entailment items (Table 1). The two
alternatives model contained only entailment items, which, under classic
accounts, should be sufficient to generate implicature, but fit to data
was quite poor with these items. The addition of a generic negative
element produced some gains, but much higher performance was found when
we included four and five alternatives, with the alternatives derived
empirically for the specific scale we used. Figure
\ref{fig:fiveAltsScatter} plots model fit for the five alternatives
model.

\begin{table}[ht]
\centering
\begin{tabular}{lll}
  \hline
model & Empirical 2-alts & Empirical 4-alts \\ 
  \hline
5-alts & r = 0.88 (alpha = 4.4) & r = 0.91 (alpha = 4.5) \\ 
  4-alts & r = 0.85 (alpha = 4.4) & r = 0.89 (alpha = 4.7) \\ 
  3-alts & r = 0.68 (alpha = 6.8) & r = 0.71 (alpha = 6.7) \\ 
  2-alts & r = 0.54 (alpha = 8.9) & r = 0.56 (alpha = 9.4) \\ 
   \hline
\end{tabular}
\caption{Model performance with fitted alpha levels. Model fit assessed through correlation with human judgments in the two conditions of Experiment 3.} 
\end{table}

@@

In our current work, we used the rational speech act framework to
investigate the set of alternatives that best allowed the model to
predict pragmatic judgments across a series of different scales. We
found that the best predictions came when a range of scale-dependent
negative and neutral alternatives were added to the implicature
computation, suggesting the importance of considering non-entailment
alternatives. This work builds on previous investigations, reinforcing
the claim that negative alternatives are critical for understanding
implicature (Franke, 2014), and replicates and extends findings that
different lexical scales produce strikingly different patterns of
inference (van Tiel et al., 2014).

@@

\hypertarget{refs}{}
\hypertarget{ref-degen2015}{}
Degen, J., \& Tanenhaus, M. K. (2015). Processing scalar implicature: A
constraint-based approach. \emph{Cognitive Science}, \emph{39}(4),
667--710.

\hypertarget{ref-frank2012}{}
Frank, M., \& Goodman, N. D. (2012). Predicting pragmatic reasoning in
language games. \emph{Science}, \emph{336}(6084), 998.

\hypertarget{ref-franke2014}{}
Franke, M. (2014). Typical use of quantifiers: A probabilistic speaker
model. In \emph{Proceedings of the 36th annual conference of the
cognitive science society} (pp. 487--492).

\hypertarget{ref-goodman2013}{}
Goodman, N. D., \& Stuhlmuller, A. (2013). Knowledge and implicature:
Modeling language understanding as social cognition. \emph{Topics in
Cognitive Science}, \emph{5}(1), 173--184.

\hypertarget{ref-grice1975}{}
Grice, H. P. (1975). Logic and conversation. In P. Cole \& J. Morgan
(Eds.), \emph{Syntax and semantics} (Vol. 3). New York: Academic Press.

\hypertarget{ref-horn1972}{}
Horn, L. R. (1972). \emph{On the semantic properties of logical
operators.} (PhD thesis). University of California, Los Angeles.

\hypertarget{ref-levinson2000}{}
Levinson, S. C. (2000). \emph{Presumptive meanings: The theory of
generalized conversational implicature}. MIT Press.

\hypertarget{ref-lewis1969}{}
Lewis, D. (1969). \emph{Convention: A philosophical study}. John Wiley
\& Sons.

\hypertarget{ref-sperber1995}{}
Sperber, D., \& Wilson, D. (1995). \emph{Relevance: Communication and
cognition} (2nd ed.). Oxford, UK: Blackwell.

\hypertarget{ref-vantiel2014}{}
van Tiel, B. J. M. (2014). Quantity matters: Implicatures, typicality,
and truth.

\hypertarget{ref-vantiel2014b}{}
van Tiel, B. J. M., Van Miltenburg, E., Zevakhina, N., \& Geurts, B.
(2014). Scalar diversity. \emph{Journal of Semantics}, ffu017.

### 2016-10-02 models/RSA_alternativeCostFormulation.Rmd
sanitize

Old RSA formulation
```{r}
# speaker informativity
# ---------------------
speaker.inf = function(d, alpha, cost = 0) {
  exp(alpha*(log(d) - cost))
}
# speaker likelihood
# ------------------
speaker.lhd = function(rating, degree, m, alpha) {
  numerator = speaker.inf(m[rating, degree], alpha)
  normalize = sum(sapply(m[rating, ], function(i) {speaker.inf(i, alpha)}))
  return(numerator / normalize)
}
# non-normalized posterior
# -----------------------
nn.post = function(rating, degree, m, alpha, useprior) {
  prior = priors[rating, "prior.p"]
  return(speaker.lhd(rating, degree, m, alpha) * prior)
}
# normalized posterior
# --------------------
norm.post = function(rating, degree, m, alpha, useprior) {
  nn = nn.post(rating, degree, m, alpha, useprior)
  normalize = sum(unlist(sapply(seq(1, 5),
                                function(i){nn.post(i, degree, m, alpha, useprior)})))
  return(nn / normalize)
}
```

New Prior incorporation
```{r}
speaker.inf3 = function(d, alpha, cost = 0) {
  if (d == 0) return(0)
  exp(-1 * alpha * (-log(d) - cost))
  # exp(alpha * (log(d) - cost))
}
# speaker likelihood
# ------------------
speaker.lhd3 = function(currRating, currDegree, m, alpha, costs, allDegrees) {
  currCost <- subset(costs, degree == currDegree)$cost
  numerator = speaker.inf3(m[[currDegree]][currRating], alpha, currCost)
  # normalize over other words
  normalize <- sum(unlist(sapply(seq(allDegrees), FUN = 
                            function(i) {
                              speaker.inf3(m[[allDegrees[i]]][currRating], alpha,
                                          subset(costs, degree == allDegrees[i])$cost)}
                          )))
  if (normalize == 0) return(0)
  return(numerator / normalize)
  # return(numerator)
}

>>>>

Old RSA formulation
```{r}
# speaker informativity
# ---------------------
speaker.inf = function(d, alpha, cost = 0) {
  exp(alpha*(log(d) - cost))
}
# speaker likelihood
# ------------------
speaker.lhd = function(rating, degree, m, alpha) {
  numerator = speaker.inf(m[rating, degree], alpha)
  return(ifelse(numerator < 0, 0, numerator))
#   normalize = sum(sapply(m[rating, ], function(i) {speaker.inf(i, alpha)}))
#   return(numerator / normalize)
}
# non-normalized posterior
# -----------------------
nn.post = function(rating, degree, m, alpha, useprior) {
  prior = priors[rating, "prior.p"]
  return(speaker.lhd(rating, degree, m, alpha) * prior)
}
# normalized posterior
# --------------------
norm.post = function(rating, degree, m, alpha, useprior) {
  nn = nn.post(rating, degree, m, alpha, useprior)
  normalize = sum(unlist(sapply(seq(1, 5),
                                function(i){nn.post(i, degree, m, alpha, useprior)})))
  return(nn / normalize)
}
```

New Prior incorporation
```{r}
speaker.inf3 = function(d, alpha, cost = 0) {
  if (d == 0) return(0)
  exp(-1 * alpha * (-log(d) - cost))
  # exp(alpha * (log(d) - cost))
}
# speaker likelihood
# ------------------
speaker.lhd3 = function(currRating, currDegree, m, alpha, costs, allDegrees) {
  currCost <- subset(costs, degree == currDegree)$cost
  numerator = speaker.inf3(m[[currDegree]][currRating], alpha, currCost)
  # normalize over other words
  normalize <- sum(unlist(sapply(seq(allDegrees), FUN = 
                            function(i) {
                              speaker.inf3(m[[allDegrees[i]]][currRating], alpha,
                                          subset(costs, degree == allDegrees[i])$cost)}
                          )))
  # return(ifelse(numerator < 0, 0, numerator))
  if (normalize == 0) return(0)
  return(numerator / normalize)
  # return(numerator)
}
